---
title: "p8105_hw_2_JL6047"
author: "Eric Luo"
date: "2022-10-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem1
First we load the data from local files, janitor::clean it and convert Entry to a logical variable
and specify that column 8-11 should be character in consistency with col 1-7

```{r}
library(tidyverse)
subway_data = read_csv(file = "C:/Users/scorp/Desktop/Mailman Year 2/8105/p8105_hw2_JL6047/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
                       col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c"))%>% 
janitor::clean_names() %>% 
select(line, station_name, station_latitude, starts_with("route"), entry, exit_only, vending, entrance_type,ada)%>%
mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
select(subway_data,entry)
skimr::skim(subway_data)
```

However, at this point the data are not tidy at this moment, and we will fix it later.

distinct stations, we use distinct() and obtain 465 unique stations
```{r}
distinct(select(subway_data,station_name,line))
```

ADA complaint unique stations: 84
```{r}
subway_data %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

Proportion: 0.377
```{r}
subway_data %>%
  filter(vending=="NO")%>%
  pull(entry)%>%
  mean
```

Distincat stations for A
```{r}
subway_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct
```
ADA compliant stations for A
```{r}
subway_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada==TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

#Problem2

Read Mr.Trash Wheel datam, create variable names, clean and round

```{r}
library(readxl)
TW_data = read_excel("C:/Users/scorp/Desktop/Mailman Year 2/8105/p8105_hw2_JL6047/Trash Wheel Collection Data.xlsx",sheet = "Mr. Trash Wheel",
col_names = c(col_names = 'dump','mon','year','date','weight','vol','PB','Pol','Cig','GlassB','GroB','CB','SB','HP','',''))[2:547,1:14]%>%
  mutate(SB= as.integer(SB))
```

Do the same thing for Professor Trash Wheel and at the same time at a line of Sports balls for reference of combining
```{r}
PT_data = read_excel("C:/Users/scorp/Desktop/Mailman Year 2/8105/p8105_hw2_JL6047/Trash Wheel Collection Data.xlsx",sheet= "Professor Trash Wheel",
col_names = c(col_names = 'dump','mon','year','date','weight','vol','PB','Pol','Cig','GlassB','GroB','CB','HP'))[2:95,1:13]%>%
mutate(SB=0)%>%
  arrange('dump','mon','year','date','weight','vol','PB','Pol','Cig','GlassB','GroB','CB','HP','SB')
```

Add a line of wheel type and merge
```{r}
TW_data = mutate(TW_data,Wheel="Mr.Trash")
PT_data = mutate (PT_data, Wheel="Professor")
  Big_data = full_join(TW_data,PT_data)
```
Mr.Trash Wheel contains 546 observations of each wheel with a total weight of 1748.36 tons of trash with 22344 homes served 
and 6879 sports balls collected

While Professor Trash Wheel contains 94 observations of each wheel with a total weight of 190.12 tons with 2709 homes served

The final dataset contains a total of 640 wheels for both Wheels with 25053 homes served

#Problem3
upload pols_month and clean it and make corresponding changes to the columns
```{r}
library(tidyverse)
pols_month = read_csv(file="C:/Users/scorp/Desktop/Mailman Year 2/8105/p8105_hw2_JL6047/pols-month.csv")%>%
janitor:: clean_names()%>%
  separate(mon, c("year","month","day")) %>%
  mutate(president = ifelse(prez_gop == "0", "dem","gop")) %>% 
  select(-prez_dem, -prez_gop,-day)
```
similar applications to snp
```{r}
snp = read_csv(file="C:/Users/scorp/Desktop/Mailman Year 2/8105/p8105_hw2_JL6047/snp.csv")%>%
  janitor:: clean_names()%>%
  mutate(date = lubridate::parse_date_time2(date,orders ="mdy", cutoff_2000 = 23)) %>% 
  separate(date, c("year","month","day")) %>%
  select(-day)%>%
  arrange(year,month)
```
upload unemployment and make changes by pivot longer
```{r}
unemployment = read_csv(file="C:/Users/scorp/Desktop/Mailman Year 2/8105/p8105_hw2_JL6047/unemployment.csv")%>% 
    pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment_rate")%>%
    janitor:: clean_names()%>%
```
Createthe final merge file and merge
```{r}
final = left_join(pols_month,snp, by = c("year","month"))
  left_join(final,unemployment,by= c("year","month"))
```

